---
title: "ggplot worksheet"
author: "Prashant Kalvapalle"
date: "2024-02-06"
output: html_document
editor: visual
---

Use the prompts in this script to understand and practice ggplot commands to make and modify plots. *If you are done with the current activity early, explore modifying the plots such as changing the colours, adding more informative annotations such as labels / spelling out the data with text.*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) # load tidyverse library : brings along ggplot with it :)
```

## Load datasets and libraries

Load the `tidyverse` library *(which loads `ggplot` as well)*

Load data-sets - [histdata](../assets/histdata.csv) and [pointsdata](../assets/pointsdata.csv) from `.csv` files using `read_csv()` command

```{r dataload}
histdata <- read.csv('../data/histdata.csv')

pointsdata <- read.csv('../data/pointsdata.csv')

```

## Histogram and density plots

Let us practice making one dimensional plots. ggplot takes only one variable, the frequency of this variable for the histogram is automatically calculated

-   Make a histogram for the variable `x1` using command `geom_histogram()` after setting up `ggplot(data =, mapping = aes(..)`

    ```{r hist1}
    ggplot(data = histdata,
           mapping = aes(x = x1)) + 
      geom_histogram()
    ```

-   *change the variable of the histogram:* Make a histogram for the variable `y1` now

    ```{r hist2}

    ```

-   Separate multiple subsets of data by colouring them with `fill = column_name`.

    -   *This will split the plot into multiple independent histograms!*

```{r hist_colour}
ggplot(data = histdata,
       mapping = aes(x = x1)) + 
  geom_histogram(aes(fill = data_colour))
```

-   Add a smoothened `geom_density` onto the histogram

```{r hist_density}
ggplot(data = histdata,
       mapping = aes(x = x1, fill = data_colour)) + 
  geom_histogram() + 
  geom_density(alpha = 0.3)
```

## Better then barplots

Now we practice plotting 2 dimensional data with a different dataset: `histdata` where

-   `x = categorical variable` : categorical =\> `factor` \~ ordered `character` vector with `levels`

-   `y = numerical variable`

1.  Draw the plot with `geom_point()`

```{r}
ggplot(histdata,
       aes(x = x1, y = data_colour)) + 
  geom_point()

```

1.  Calculate mean and plot

```{r}
hist2 <- mutate(histdata, .by = data_colour,
                mean_x1 = mean(x1))

```

```{r}
ggplot(hist2,
       aes(x = x1, y = data_colour)) + 
  geom_point() + 
  geom_point(aes(x = mean_x1), 
             colour = 'red', shape = '|', size = 5)
```

1.  Add text label on the mean
2.  Connect each replicate across the conditions with a `geom_line()`
3.  Label the individual replicates

## Points and lines

Now we practice plotting 2 dimensional data where both x and y are numerical variables

1.  Make y1 vs x1 plot with points

    ```{r points}

    ```

2.  Join the points with a line

    ```{r addlines}

    ```

3.  Add colours for points

    ```{r addcolours}

    ```

4.  Attempt something fancy

-   Add labels

-   Show mean on the graph etc..
